### 2022-10-04 Proposal/Proposal.tex
commit #setupproposal

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\title{My Statistics 3494W Proposal}

\author{Michael  Marcaccio}

\date{October 4 2022}

\begin{document}

\end{document}