### 2022-10-04 Proposal/Proposal.tex
commit #firstcite

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\title{My Statistics 3494W Proposal}

\author{Michael  Marcaccio}

\date{October 4 2022}

\begin{document}

\maketitle

\section*{Introduction}
Call centers are a staple of the finanical industry with agents working in the United States, Europe, Asia and Africa \citep{ibrahim2016modeling}.

\bibliographystyle{chicago}
\bibliography{citations.bib}
\end{document}