### 2022-10-04 Proposal/Proposal.tex
commit #Introduction

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\title{My Statistics 3494W Proposal}

\author{Michael  Marcaccio}

\date{October 4 2022}

\begin{document}

\maketitle

\section*{Introduction}
Call centers are a staple of the finanical industry with agents working in the United States, Europe, Asia and Africa \citep{ibrahim2016modeling}. 
Effective management is essential to running call centers and businesses are consistently trying to model and optimize when customers are calling with questions. 
\citet*{evensen1999effective} describes the recent growth of call centers as being, "viewed as little more than lower cost channels for customer
problem resolution, are quickly becoming a powerful means of service delivery with a potential for substantial revenue generation".
I have chosen this to be my topic of study as I have a strong interest in working in the finanical industry after graduation and I have an unique opportunity to utilize real-world data from an undisclosed
finanical firm. Due to this, My findings may provide insight that would benefit the company and its members.

\section*{Specific Aims}
I will be attempting to model call volume per a 15 minute interval and determining if there is any seasonality in the calls.

\bibliographystyle{chicago}
\bibliography{citations.bib}
\end{document}