### 2022-11-12 Manuscript/Manuscript.tex
commit #starting correlation

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{tabularx} 
\usepackage{enumitem}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\usepackage{underscore}
\MakeOuterQuote{"}
\title{Call Center Regression Data Analysis}

\author{Michael  Marcaccio}

\date{November 14 2022}

\begin{document}

\maketitle

\begin{abstract}
 With over 15 million people employed, a predicted compound annual growth rate (CAGR) of 5.6 percent between 2020 and 2027, and with over 28,000 locations
 in the United States alone, call centers play a pivotal role in a business' success.In this paper, call center volume was forecasted and
 models were created to predict the number of agents needed to meet critical attributes such as waittime, calltime, and holdtime. 
 Forecasting gives businesses the ability to make informed business decision and develop data-driven strategies. In the literature it is very
 common to see call center volume being predicted using different techniques, but there is very limited studies on attempting to correlate
 the number of agents needed. Through Regression techniques, there are 8 models proposed to model the number of agents needed based on waittime,
 calltime, goaltime, and the amount of calls handled.
\end{abstract}

\section*{Introduction}
  With over 15 million people employed, a predicted compound annual growth rate (CAGR) of 5.6 percent between 2020 and 2027, and with over 28,000 locations
in the United States alone, call centers play a pivotal role in a business' success. Call centers are a key part of customer service
that will save a company time, money, and unneccessary obstacles. In the banking industry, calls can range from inquires, transfers, 
payments, reporting, to processing. This means members could call about their account balance, credit card bills, loan applications, 
or unauthorized transactions. It is crucial that a bank is prepared for spikes in calls and have agents knowledgeable in all aspects of banking.

  Many studies have been done working with call center volumes as presented in Modeling and Forecasting Call Center Arrivals \citep{ibrahim2016modeling}.
Here an Autoregressive moving average (ARIMA) model is used, which is depicted in the paper as:
\begin{equation}
  \label{eq:ARIMA Standard}
  \Phi(B)(x_i -\mu)=\theta_q(B)\varepsilon_t
\end{equation}
Ibrahim uses multiple ARIMA models to depict seasonality with a combination of exponential smoothing. Holt-Winters smoothing is also used,
with three equations of:
\begin{equation}
  \label{eq:Holt-Winters}
  M_t=\alpha_0(X_t-S_{t-s}) + (1-\alpha_0)(M_{t-1}+B_{t-1}),
\end{equation}
\begin{equation}
  \label{eq:Holt-Winters2}
  B_t=\alpha_1(M_t-M_{t-1}) +(1-\alpha_1)B_{t-1},
\end{equation}
\begin{equation}
  \label{eq:Holt-Winters3}
  S_t=\alpha_2(X_t-M_t) +(1-\alpha_2)S_{t-s},
\end{equation}
where $B_{t}$ is the slope component, $M_{t}$ is the level component, $S_{t}$ is the seasonal component, and s is the period of seasonality.
ARIMA models are a great model to create as "they allow the  representation of a  wide array of potentially useful predictor functions in  models  which contain relatively few parameters" \citep{newbold1983arima}.
With the combination of Holt-Winters smoothing, Ibrahim creates a great model for forecasting call volume, however the paper never goes into detail about
how many agents there should be working at a set time.

  \citep{evensen1999effective} goes into detail about effective service delivery stating the expectations of customers are a "function of
the customer's own experiences...and when judging their own service quality, financial instiutions need to evalute themselves on 
objective measures which span across industries".When taking this into consideration, I focused my models on using a calls per agent approach,
so agents are not overworked and can treat each caller with respect while answering their concerns in a timely manner to represent the business in a good light. \citep{avramidis2005modeling} confirms
this as it is described as the "call-to-agent assignment problem", and that an efficiency-driven call center is important.

  The contributions of the paper creates models on the basis to maximize quality of service in relation to the number of agents. 
Based on this, models were created to minimize:
\begin{enumerate}
  \item Wait Time
  \item Call Time
  \item Reduce the Amount of Agent Handle Time
  \item Reduce the Amount of Calls not being handled
\end{enumerate}
The remainder of the paper will present the dataset and break it into a clear view on the historical data. I have also created a 
forecast on the projected call volume throughout a week on 15 minute intervals.

\section*{Data}
\begin{sloppypar}
  The data comes from an undisclosed bank where I wrote two SQL queries to obtain. This comes directly from the banks records dating back from Febuary 2020
  to Septemeber 2022, and I was lucky enough to do a data analysis on real-life data. This dataset consists of 466,565 observations of 31 variables. 
  Some significant variables consist of Calldate, Caller_ID, Caller_ID, QueuedYN, AnsweredYN, AbandonedYN, QueuedName, InteractionOutcome, 
  AnsweredGoalMet, StartTime, EndTime, WaitTime, Holdtime, AgentTalkDuration, AgentHandleTime, WrapUpTime, DayoftheYear,
  WeekoftheYear, and NbrofAgents.
\end{sloppypar}
  It is important to acknowledge that AgentTalkDuration refers to the actual amount of time the agent the agent was talking while
AbandonedYN refers to if the caller ended the call before speaking to a representative. The QueuedYN is if the member was put in a queue and had to wait to speak to an agent.
The queue is an automated service or an IVR (Interactive Voice Response). The latest generation of speech-recognition technology allows
IVRs to interpret complex user commands, so customers may be able to “self-serve”, i.e., complete the service interaction at the IVR. \citep{avramidis2005modeling}.
Holdtime is how long the member waits when an agent places on them hold, while waittime is how long they have to wait before someone answers there call.
WrapUpTime refers to how long the member takes to hang up the phone after the agent stops talk. AgentHandleTime is the total of Holdtime, WrapUpTime, and AgentTalkDuration combined.

  This data is allows me to easily answer the research question as I am using real-life data. With a huge sample size and many variables,
this is a great opportunity for data analysis and modeling. With this dataset, I know when doing calculations such as finding means and creating plots, 
I will able to accurately give a truly image of call center volume and agent information to a small period of time such as 15 minutes.
  I did do some exploratory data analysis by creating figures that represent call center volume by the year, month, day of week, and time. 
I also created a tabled that shows the average wait time, talk time, hold time, wrap up time, and count of every Interaction Outcome.
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{By Year.png}
  \caption{Call Volume Grouped by Year.}
  \label{fig:Year}
\end{figure}
Here we can see that 2021 has the highest call volume. Please Note that 2020 only has 11 months, while 2020 only has 10. This may be why the data appears skewed.
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{ByMonthVolume.png}
  \caption{Call Volume Grouped by Month.}
  \label{fig:Month}
\end{figure}
When grouped by month, we can see an interesting peak around April. This may be because that is tax season and members are calling for tax information.
Also April 2020 may have a higher peak than April 2021, due to COVID-19. During that time, many individuals were eligble for a 1,200 dollar stimulus check.
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{CallVolumeByDayOfWeek.png}
  \caption{Call Volume Grouped by Day.}
  \label{fig:Day}
\end{figure}
With Sunday representing 1, we can see that most calls happen on a Friday. Saturday has an extremely low count. This can be explained due to the
varying hours of the call center. Monday through Thursday, the call center is open 8am to 5pm, a total of 9 hours. Friday the call center is open
8am-7pm, a total of 11 hours. Saturday the call center is only open 8:30 am to 12pm, a total of 3 and a half hours.
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{CallVolumeByTimeOfDayGROUPEDbyYear.png}
  \caption{Call Volume Grouped by Time.}
  \label{fig:Timeoneday}
\end{figure}
The call center consistently has a spike around 9am throughout the 3 years. This is suprising as the call center opens at 8am. This is important to be aware of when
creating a forecasting model and having the banks employees be at their station at 9am. The decreasing trend throughout the day is not suprising
as most people are working throughout the day.
\begin{table}[H]
  \resizebox{\textwidth}{!} {
  \begin{tabular}{ l | l | l | l | l | l | l}
    {\bf Outcome} & {\bf Count} & {\bf Wait Avg} & {\bf Talk Avg} & {\bf Hold Avg} & {\bf Wrap Avg} & {\bf Total Talk Avg} \\
  \hline
  Abandoned & 2364 & 100 & 11.3 & 0.06 & 2.40 & 13.8 \\
  \hline
  Handled & 460047 & 107 & 253 & 26.2 & 16.3 & 295\\
  \hline
  Leave Number & 411 & 214 & 66 & 6.18 & 7.80 & 80\\
  \hline
  Not Specified & 299 & 1.21 & 120 & 0.278 & 3.24 & 123\\
  \hline
  Service Unavailable & 1 & 0 & 0 & 0 & 0 & 0\\
  \hline
  Wrong Number & 3 & 357 & 0 & 0 & 12 & 12\\
  \end{tabular}
  }
  \end{table}
Here it is prevalent that a majority of calls are handled. These variables played a key part in my models for the agents. It is also to see that 
100 seconds is the cutoff for most people that Abandoned. That could be a goal for this bank to get too, to prevent more Abandoned calls. 
\section*{Methods}
When seeing these patterns, I did not see much seasonality in the data. While there was a clear jump in volume based on time, I did not see much in regards
to year, month, or day. I then confirmed this by making one last histogram.
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{SeasonalityOfCallVolume.png}
  \caption{Seasonality Of Call Volume .}
  \label{fig:Seasonality}
\end{figure}
Here we do not see any clear pattern based on whether its winter, spring, summer, or fall. Overall, this was not too surpising.

I also created linear models to show the relationship between calls per agent and average waitime, holdtime, average goal met, and average 
interaction value. Average goal met was a variable created in relationship to the AnsweredGoalMet. AnsweredGoalMet was given in 1 and 0s, 
1 if the goal was hit and 0 if it was not hit. This variable is very fluid, as each company has its own reasoning for what they want their
goal time to be. Average Interaction Value was based off a variable that was created based on the InteractionOutcome. "Handled "got assigned 
1 as it was a positive outcome and -1 for "Abandoned" and "Leave Number" as they were deemed negative. The others were considered neutral, for example
"Not Specified", was given a 0.

  Linear Regression is based off five key assumptions,
  \begin{enumerate}
  \item There is a Linear relationship,
  \item Multivariate normality,
  \item No or Little multicollinearity,
  \item No auto-correlation,
  \item Homoscedasticity
\end{enumerate} Multicollinearity refers to the occurrence of high intercorrelations among two or more independent variables in a multiple regression model.
To prevent this, I soley focus on one variable at a time when making my models on agents. Homoscedasticity is an assumption there is equal or similar variances in different groups being compared.
From here I then did a Correlation Plot:
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Correlation.png}
  \caption{Correlation Between Variables.}
  \label{fig:Correlation}
\end{figure}
Here we see that correlations between Average Wait Time, Average Handle Time, Average Answer Goal Time, Average Score Answer Goal Time Met, and Ratio Calls Per Agent.
We see that Calls Per Agent is correlated with all 4 variables in our models which was a great step to acheive. I then checked to see what the distribution was between calls per agent was throughout each day. 
  \begin{table}[H]
    \resizebox{\textwidth}{!} {
    \begin{tabular}{ l | l | l | l | l |}
      {\bf DayofWeek} & {\bf AvgNumCalls} & {\bf AvgNumAgents} & {\bf CallsPerAgent} & {\bf Num15Periods} \\
    \hline
    2 & 3.17 & 10.3 & 0.308 & 38 \\
    \hline
    3 & 3.25 & 9.81 & 0.331 & 39 \\
    \hline
    4 & 3.17 & 10.4 & 0.305 & 38 \\
    \hline
    5 & 3.08 & 10.1 & 0.306 & 37 \\
    \hline
    6 & 3.83 & 8.82 & 0.435 & 46 \\
    \hline
    7 & 1.25 & 6.33  & 0.197 & 15 \\
    \end{tabular}
    }
    \end{table}
Here we see that the Calls Per Agent is highest on Friday. This further proves that Friday is the busiest. Note that these are on 15 minute intervals,
so for example, the call center gets 3.17 calls every 15 minutes on average on a Monday. It was surpising to see that Friday was the second
lowest staffed day of the week. A possible suggestion to the company would be add an agent on Friday. Members may be calling about last minute
questions on Friday to enjoy the weekend and not have to call Saturday. In regards to Saturday, it is clear that is the least busiest day of the week,
this is not too suprising as many people like to sleep in on the weekends and not worry about financial issues. Monday through Thursday is relatively
consistant in all aspects. Note that the reason why the "num15minPeriods" variable is not the same is due to the outliers in the dataset where individuals
called either slightly before or after call center hours.
\section*{Simulation}
After seeing how consistent the call volume was and how big of a sample size I had, I decided to create a table and go off total calls and use averages to project call volume per weekday with the 
created variable called RatioDayPerWeek. This variable is the ratio of how many hours the call center was open historically on the day(Monday,Tuesday,Wednesday...) compared to all the total hours of all days.
I was able to produce this table:
\begin{table}[H]
  \resizebox{\textwidth}{!} {
  \begin{tabular}{ l | l | l | l | l |}
    {\bf DayofWeek} & {\bf numberofCalls} & {\bf AvgCallsPerDay} & {\bf RatioDayPerWeek} & {\bf ProjectedCallsPerDay} \\
  \hline
  2 & 79785 & 670 & 0.193 & 642 \\
  \hline
  3 & 89121 & 655 & 0.188 & 628 \\
  \hline
  4 & 89428 & 653 & 0.188 & 625 \\
  \hline
  5 & 84918 & 624 & 0.180 & 598 \\
  \hline
  6 & 95588 & 693 & 0.199 & 664 \\
  \hline
  7 & 24285 & 183 & 0.0525 & 175\\
  \end{tabular}
  }
  \end{table}
This table shows that Friday accounts for 19.9 percent of all the hours in the week the call center is open, while Saturday only accounts
5.25 percent of the hours. The only difference there is a different in ratio for Monday thru Thursday must be holidays, e.g. Thanksgiving. 
The Projected Calls Per Day simply came from the total calls per week * the RatioDayPerWeek. We can also see this approach worked relatively well in the forecasting model. 
We also see that the MAPE(Mean Absolute Percent Error) is 4.21 percent. For context, a MAPE score less than 5 is very accurate.

  Using this information, I was able to create a figure that showed the amount of expected calls per day per 15 minute intervals. 
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Forecast15MinPeriods.png}
    \caption{This is the Forecast Per Day On a 15 Minute Intervals.}
    \label{fig:Forecast}
  \end{figure}
Here we see that distinct pattern similar to the historical graph. The graph peaks at 9am and then slowly trails throughout the day.
Note that the different colors represent the different days.

In regards to forecasting the number of agents, I create one last table to see if there is a difference in AgentHandleTime and WaitTime
by each 15 minute interval.
\begin{table}[H]
  \resizebox{\textwidth}{!} {
  \begin{tabular}{ l | l | l | l | l | l | l |}
    {\bf DayofWeek} & {\bf AvgNumCalls} & {\bf AvgNumAgents} & {\bf CallsPerAgent} & {\bf AvgAgentHandleTime} & {\bf AvgWaittime} & {\bf AvgInteractionValue}\\
  \hline
  2 & 2418 & 10.6 & 227 & 303 & 175 & 0.987 \\
  \hline
  3 & 2701 & 10.4 & 260 & 297 & 126 7 0.998\\
  \hline
  4 & 2710 & 10.7 & 254 & 297 & 80.9 & 0.987 \\
  \hline
  5 & 2573 & 10.2 & 253 & 289 & 87.1 & 0.988 \\
  \hline
  6 & 2897 & 9.93 & 292 & 293 & 88.2 & 0.987 \\
  \hline
  7 & 736 & 6.34 & 116 & 260 & 57.4 & 0.989\\
  \end{tabular}
  }
  \end{table}
Overall, we see that Friday continues to be the busiest days in terms of call volume but the least staffed days. However, a glaring note
is how high the average waittime on Monday is. This may be suggesting that the subject matter of calls on Monday are more intense and urgent
then the subject matter on a day like Saturday. With such a high wait time but not a bigger handle time, this also may be suggesting that
employees are getting stressed out at lot more and may need more time to get themselves mentality ready to take the next call. It might be a good idea to add another employee for Mondays. 
The other days appear to be relatively normal and Saturday is functioning very efficiently. With a waittime less than 60 seconds, there is no reason to change the staffing model for
Saturday.

Next were Correlation tests

\section*{Discussion}

\bibliographystyle{chicago}
\bibliography{Citations.bib}
\end{document}